---
title: "Cluster Analysis outline"
author: "Victor Mandela"
date: "January 7, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
 
```{r}
library(tidyverse)
theme_set(theme_light())
```

## Defination

Clustering is a form of exploratory data analysis (EDA) where observations are divided into meaningful groups that share common characteristics (features).

## The flow of cluster analysis

1. Pre-processing

   - no missing values
   
   - features in similar scale
   
2. Select similarity measures

3. Clustering method

4. Analyse

Market segmentation and pattern grouping are good examples where cluster analysis applies.

## Examples

```{r}
# Two players

two_players <- as.data.frame(list(x=c(0,9),y=c(0,12)))

#plot the the position of the two players

ggplot(two_players,aes(x=x,y=y)) + geom_point(size =4, shape = 7) + scale_x_continuous(limits = c(0,12))
```

Distance = 1 - similarity

We use the euclidean distance in this case to find the distance.

Note: The larger the distance indicates the dissimilarity between observations.

### Three players

With more than two observations it can be tiresome calculating the euclidean distance manually, thats why we introduce the 'dist()'
 function. The method in this case would be 'euclidean'.
 
```{r}
three_players <- as.data.frame(list(x = c(0,9,-2), y = c(0,12,19)))

rownames(three_players) <- c("Ronaldo", "Gabriel", "Messi")

#plot the players
ggplot(three_players,aes(x = x, y = y)) + geom_point(size = 4) + scale_x_continuous(limits = c(-4,10)) + scale_y_continuous(limits = c(0,20))

```

In this case we will use the 'dist()' function.

```{r}
# Find the distance
dist(three_players)

```

## Different scales

The importance of scaling the data is to *standardize* our feature to a mean of zero and a standard deviation of 1. (Normalizing the data!!!)



## Example2

We are going to use the 'Nutrient' dataset from the 'flexclust' package.

```{r}
library(flexclust)
data("nutrient");nutrient
```

Since the data has different measurement units we need to scale.

```{r}
scale(nutrient)
```

## Hierarchical clustering

We compare more than two observation. 

```{r}
dist_nutrient <- dist(nutrient, method = "euclidean")
hc_nutrient <- hclust(dist_nutrient, method = "complete")
```

# Extract K clusters

```{r}

clusters_assignment <- cutree(hc_nutrient,k = 2)

#Create a  new column
nutrient_clustered <- mutate(nutrient, cluster = clusters_assignment)

#visualize the clusters
ggplot(nutrient_clustered, aes(x = protein, y = iron, color = factor(cluster))) + geom_point()
```


